Online Laterality Battery Pre-Processing - short version

### 2020-05-24 battery-preprocessing_v5_DB.Rmd
Finalised RDT analysis - confirms this is no good! But finger tapping is much better than we thought: I had confused the labels of columns for this, so L/R was mixed up with day1/day2. Now it is fixed, this gives a nice laterality index which is strongly related to handedness and which has reliability just above .7

This portion of the notebook prcoesses the "Rhyme Detection Task' implemented online via Gorilla.sc, which aims to assess the lateralisation of language processing. This is a new task devised for this project. In the task, participants are shown a word in the central visual field and two images in the left and right visual field. The task is to judge which of the two images has a name that rhymes with the central word.

@@

```{r RDT.LI}
#now create LI
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"RDT_p.corr_LI"
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"RDT_RT_LI"
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)

#Cutoff has removed those with low accuracy, but we should also remove their RT LIs
#We do this by finding those with NA in accuracy LI
w<-which(is.na(allsum$RDT_acc_LI))
if(length(w)>0)
{allsum$RDT_RT_LI[w]<-NA}
#t.test to test if laterality index differs from zero
partsum <- allsum[allsum$exRDT==0,]
print('Testing for difference of single mean from zero')
t.test(partsum$Day1.RDT_p.corr_LI)
t.test(partsum$Day1.RDT_RT_LI)
t.test(partsum$Day2.RDT_p.corr_LI)
t.test(partsum$Day2.RDT_RT_LI)
#rename column to identify day
myncol <- length(colnames(allsum))
colnames(allsum)[(myncol-1):myncol] <- paste0('Day',d,'.',colnames(allsum)[(myncol-1):myncol])

}
cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

@@

So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. Alas, the RDT task has to be ditched!

@@

This portion of the R Markdown file prcoesses the "Finger Tapping Task' implmented in Gorilla.sc which aims to quantify handedness. In the task, participants are required to press a sequence of buttons twice for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within a 30 ms. There is a version that requires more precise movements.

A score of 1 is given each time that a participants completes the sequence.

This reads in a csv file where each row represents a button press. Participants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

@@

```{r readdataFinger, warning=FALSE}

finger <- read.csv("day1_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

all_finger1<- finger

finger <- read.csv("day2_combined.csv", na.strings = "NA")

# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task
# write as factors
finger$display <- as.factor(finger$display) 
finger$subject <- as.factor(finger$Participant.Public.ID)
finger$RT <- as.numeric(as.character(finger$Reaction.Time))

@@

@@

subids <- levels(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject', 'R1', 'R2', 'L1', 'L2', 'R1sd', 'R2sd', 'L1sd', 'L2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #will be incremented so sums written to correct columns
  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
      tempfinger <- filter(myfinger,subject==subids[n],Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x')) 
      #exclude rows with other stuff in Response
      #(this is not essential as nonsequences will just get ignored)

      wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
      for (s in 1:2){
        myc<-myc+1 #increment column counter
        target <- myseq[s]
        finger.df[n,myc]<-str_count(wholeseq,target) 
        #count how many target sequences are in the long string that contains all responses

      }
      righth<-filter(tempfinger,display=='right hand')
      lefth<-filter(tempfinger,display=='left hand')
      startcol<-6
      if(d==2){startcol<-8}
      finger.df[n,startcol]<-sd(righth$RTa,na.rm=T)
      finger.df[n,(startcol+1)]<-sd(lefth$RTa,na.rm=T)
    }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

@@

```{r domakeZ_finger}
#We add Chimeric faces LIs to allsum for day 1 and then day2
task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (-1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z)
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy')
abline(v=0)
abline(h=0)
text(-2.5,2.5,paste0('r = ',round(cor1,3)))

@@

```{r corr_between_tasks}
mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','tlat_RDT_1','tlat_RDT_2','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
colnums<- which(colnames(allsum) %in% mycols)
mycorrs<-correlate(allsum[,colnums],method='spearman')

>>>>

This portion of the notebook prcoesses the "Rhyme Detection Task' implemented online via Gorilla.sc, which aims to assess the lateralisation of language processing. This is a new task devised for this project. In the task, participants are shown a word in the central visual field and two images in the left and right visual field. The task is to judge which of the two images has a name that rhymes with the central word. The logic of the task is that to perform one must generate the phonology of a pictured word's name, which is a classic left-hemisphere task - it is assumed this will be easier for pictures the are projected to the left hemisphere (ie RVF). 
A small proportion of trials are 'no rhyme' cases, which are designed to assess attention to the task.

@@

```{r RDT.LI}
#now create LI
for (d in 1:2){
origdat <-RDT_df1[RDT_df1$subject %in% incRDT,]
if (d==2)
{origdat <-RDT_df2[RDT_df2$subject %in% incRDT,]}
origdat$X1 <- origdat$p.corr
origdat$side <- as.factor(origdat$side)
levels(origdat$side) <-c('Left','No rhyme','Right') #function assumes capitalised
mycolname<-"Day1.RDT_p.corr_LI"
if(d==2){mycolname<-"Day2.RDT_p.corr_LI"}
polarity<- 1 #so LI is positive in predicted direction
mycutoff <- 75
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
origdat$X1 <- origdat$RT
mycolname<-"Day1.RDT_RT_LI"
if(d==2){mycolname<-"Day2.RDT_RT_LI"}
polarity <- (-1)
allsum <- makeLI(allsum,origdat,task,mycolname,polarity,mycutoff)
}

#t.test to test if laterality index differs from zero
partsum <- allsum[allsum$exRDT==0,]
print('Testing for difference of single mean from zero')
t.test(partsum$Day1.RDT_p.corr_LI)
t.test(partsum$Day1.RDT_RT_LI)
t.test(partsum$Day2.RDT_p.corr_LI)
t.test(partsum$Day2.RDT_RT_LI)

cor1<-cor(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,use='complete.obs')
cor2<-cor(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,use='complete.obs')
plot(partsum$Day1.RDT_p.corr_LI,partsum$Day2.RDT_p.corr_LI,main='Rhyme detect: LI for accuracy (%correct)')
abline(v=0)
abline(h=0)
text(-3,3,paste0('r = ',round(cor1,3)))
plot(partsum$Day1.RDT_RT_LI,partsum$Day2.RDT_RT_LI,main='Rhyme detect: LI for RT')
abline(v=0)
abline(h=0)
text(-3,9,paste0('r = ',round(cor2,3)))
```
## Does RDT relate to handedness?
Answer is convincing nope!

@@

So the paired items analysis gives a totally nonsignificant correlation between time 1 and time 2. This is much worse than the overall LI reliability.

The failure of this task led to further thoughts about why it is barely lateralised, given that it does implicate language-specific word generation. 
A noteworthy feature of the other two tasks (CF and dichotic) is that they measure side preference rather than accuracy - i.e. a response from either L or R is correct, but two potential responses are put in competition. That contrasts with this task where only one side has the correct answer.  
We used pictures because with written words there is a concern that  preference may be confounded by the fact we read words from L to R. That could potentially dilute the RVF advantage that is expected from L hem language.

@@

This portion of the notebook processes the "Finger Tapping Task' implmented in Gorilla.sc, which aims to provide a quantified measure of hand skill. In the task, participants are required to press a sequence of buttons for each hand (Left: W, R, V, X; Right: T, U, M, B) as many times as possible within 30 ms. We have also devised is a version that requires more complex movements, but early data suggested that does not achieve our aim of emphasising lateral differences.

A score of 1 is given for each sequence completed within the time interval.

In the raw data file, farticipants, identified by random letter strings, are stacked on top of each other. We start by reading the data and cutting out unwanted information.

@@

```{r readdataFinger, warning=FALSE}
for (d in 1:2){
finger <- all_dat2[all_dat2$Task.Name=='Finger Tapping',]
if(d==1){
finger <- all_dat1[all_dat1$Task.Name=='Finger Tapping',]}
# cleaning
finger <- finger[finger$Attempt >= 1,] # only answers 
finger <- finger[finger$display == "right hand" | finger$display == "left hand",] # only for task

@@

if(d==1){all_finger1 <- finger}
}
```

@@

subids <- unique(finger1$subject) #just look at data for those doing time1 and
nsub <- length(subids)

finger.df <- data.frame(matrix(NA,nrow=nsub,ncol=9))
colnames(finger.df)<-c('subject', 'R1', 'R2', 'L1', 'L2', 'R1sd', 'R2sd', 'L1sd', 'L2sd')
for (n in 1:nsub){
  finger.df$subject[n] <- subids[n]
  myc <- 1 #column index: will be incremented so sums written to correct columns
  mycols<-matrix(c(2,4,
                 3,5),nrow=2,byrow=T)
  #makes a matrix for looking up column to write to : rows are days, cols are R then L

  for (d in 1:2){ #each day data
    myfinger <- finger1
    if(d==2){
      myfinger<- finger2}
    tempfinger <- filter(myfinger,subject==subids[n],
                         Response %in% c('t', 'u', 'm', 'b', 'w', 'r', 'v', 'x'))  
    #all rows with these keypresses for this subject

    wholeseq <-paste0(tempfinger$Response,collapse='') #make a long string of responses
    for (s in 1:2){
      myc<-mycols[d,s] #select column
      target <- myseq[s] #tumb then wrvx, ie right then left
      finger.df[n,myc]<-str_count(wholeseq,target) 
      #count how many target sequences are in the long string that contains all responses

    }
    #We then add the SD of the RT - this indicates variability in pace of keypressing
    righth<-filter(tempfinger,display=='right hand')
    lefth<-filter(tempfinger,display=='left hand')
    myc<-mycols[d,1]+4

    finger.df[n,myc]<-sd(righth$RTa,na.rm=T)
    finger.df[n,(myc+2)]<-sd(lefth$RTa,na.rm=T)
  }
}
# Substitute NA for any totals less than 5 - suggest using wrong keys perhaps

@@

```{r domakeZ_finger}

task<-'Finger'
origdat<-finger.long
origdat$X1 <- origdat$`1`
mycolname<-"Day1_Finger_acc_Z"
mycutoff <- 10 #NB! need to check against prereg. I think this is N correct, rather than percentage. People with avg scores below this value will be excluded.
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

origdat<-finger.long
origdat$X1 <- origdat$`2`
mycolname<-"Day2_Finger_acc_Z"
polarity<- (1)#so LI is positive in predicted direction
allsum <- makeZ(allsum,origdat,task,mycolname,polarity,mycutoff)

cor1 <- cor(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,use='complete.obs')
plot(allsum$Day1_Finger_acc_Z,allsum$Day2_Finger_acc_Z,main='Finger Z accuracy',col=allsum$handpch,pch=15)
abline(v=0)
abline(h=0)
text(-2.5,4,paste0('r = ',round(cor1,3)))

@@

```{r corr_between_tasks}
mycols <- c('Day1_CF_acc_Z','Day2_CF_acc_Z','Day1_Dich_acc_Z','Day2_Dich_acc_Z','Day1_Finger_acc_Z','Day2_Finger_acc_Z')
colnums<- which(colnames(allsum) %in% mycols)
mycorrs<-correlate(allsum[,colnums],method='spearman')
